---
title: "Food Choices and Preference of College Students"
author: "Charlene"
date: "2023-08-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading readr package

Dataset used for this project was imported using readr package. To do so, the tidyverse package which contains readr package was installed and loaded.

## Importing CSV file

```{r import data, echo=FALSE, message=FALSE}
food_survey_respond_df <-readr::read_csv(here::here("Data", "food_coded.csv"))
```

## Data cleaning

#### Preview dataset

Packages required in this step: Here, Skimr, Janitor, dply.

```{r head()}
head(food_survey_respond_df)
```

```{r glimpse()}
dplyr::glimpse(food_survey_respond_df)
```

```{r colnames()}
colnames(food_survey_respond_df)

```

## Data Transformation

In this part of the analysis, 3 new dataframes will be created with variables of interest that were required to answer the following questions:

-   Question 1: Does exercising regularly affects weight of an individual?

-   Question 2: Does factors such as: Employment, Income, and Living on or off campus affect the frequency of eating out for college students?

-   Question 3: Does students who exercise regularly more likely to make better food choices?

#### Question 1: Creating new dataframe with variables of interest

In the preview of the food_survey_respond_df, there are 61 variables in this dataset. Not all the variables will be needed in our analysis, therefore, only the variables of interest will be selected and stored in a new dataframe for easy access later.

```{r}
library(tidyr)
library(dplyr)
weight_df<- food_survey_respond_df %>% 
  select("Gender","exercise","weight") %>% 
  drop_na()

weight_df %>% 
  count(weight) # to identify which responds needs cleaning / removing

weight_df %>% 
  count(Gender) # should have 1 & 2 only. 1= Female, 2 = Male

weight_df %>% 
  count(exercise)
```

There are 4 rows in weight that needs cleaning before we can start analysis.

```{r}
library(stringr)
weight_df$weight <- str_replace_all(weight_df$weight, "^144 lbs$", "144")
weight_df$weight <- str_replace_all(weight_df$weight, "^Not sure, 240$", "240")

#Removing Row 3 & row 66
weight_df<- weight_df[-c(3,66),]

#Double checking the data within weight to make sure we have cleaned up properly
weight_df %>% 
  count(weight)
```

#### Question 1: Replace Integer respond for Gender with String

```{r}
weight_df <- weight_df %>% 
  mutate(Gender_new=case_match(
           Gender,
           1~'Female',
           2~'Male'
         )
         )

weight_df %>% 
  count(Gender, Gender_new)
```

#### Question 1: Replace Integer respond for exercise with String

```{r}
weight_df <- weight_df %>% 
  mutate(exercise_new=case_match(
           exercise,
           1~"Everyday",
           2~"2-3 times a week",
           3~"Once a week",
           4~"Sometimes",
           5~"Never"
         ) 
         )

weight_df %>% 
  count(exercise_new, exercise)  
```

#### Question 1: Replace Integer respond for weight with String
```{r}
weight_df <- weight_df %>% 
   mutate(weight_range = case_when(
     weight <130 ~'110 - 129 lbs',
     weight <160 ~ '130 - 159 lbs',
     weight < 190 ~'160 - 189 lbs',
     TRUE ~ '>=190 lbs'
     )
     )
weight_df %>% 
  count(weight_range, weight)
```

#### Question 2: Creating new dataframe with variables of interest

```{r new dataframe for frequency of eating out & regular exercise leads to making better food choice}
frequency_df <- food_survey_respond_df %>% 
  select ("eating_out","employment","income","on_off_campus") %>% 
  drop_na()

glimpse(frequency_df)
```

#### Question 2: Replace Integer respond for eating_out with String

```{r}
frequency_df <- frequency_df %>% 
  mutate(eating_out_new= case_match(
    eating_out,
    1 ~ "Never",
    2 ~ "1-2 times",
    3 ~ "2-3 times",
    4 ~ "3-4 times",
    5 ~ "Everyday"
  )
  )
frequency_df %>% 
  count(eating_out_new, eating_out) # checking if above code has been run correctly
```

#### Question 2: Replace Integer respond for employment with String

```{r}
frequency_df <- frequency_df %>% 
  mutate(employment_new = case_match(
    employment,
    1 ~ "Full time",
    2 ~ "Part time",
    3 ~ "Unemployed"
  )
  )
frequency_df %>% 
  count(employment_new, employment)
```

#### Question 2: Replace Integer respond for income with String

```{r}
frequency_df <- frequency_df %>% 
  mutate(income_range = case_match(
    income,
    1 ~ "less than $15K",
    2 ~ ">$15K - $30K",
    3 ~ ">$30K - $50K",
    4 ~ ">$50K - $70K",
    5 ~ ">$70K - $100K",
    6 ~ "more than $100K"
  )
  )

frequency_df %>% 
  count(income, income_range)
```

#### Question 2: Replace Integer respond for living on and off campus with String

```{r}
frequency_df <- frequency_df %>% 
  mutate(living_on_off_campus = case_when(
    on_off_campus == 1 ~ "On campus",
    TRUE ~ "Off campus"
  )
  ) 

frequency_df %>% 
  count(on_off_campus, living_on_off_campus)

```

#### Question 3: Creating new dataframe with variables of interest

```{r}
exercise_df <- food_survey_respond_df %>% 
  select ("exercise","nutritional_check","veggies_day","fruit_day") %>% 
  drop_na()

glimpse(exercise_df)
```

#### Question 3: Replace Integer respond for exercise with String

```{r}
exercise_df <- exercise_df %>% 
  mutate(exercise_new = case_match(
    exercise,
    1 ~ "Everyday",
    2 ~ "2-3 times a week",
    3 ~ "Once a week",
    4 ~ "Sometimes",
    5 ~ "Never"
  )
  ) 

exercise_df %>% 
  count(exercise, exercise_new)

```

#### Question 3: Replace Integer respond for Nutritional check with String

```{r}
exercise_df <- exercise_df %>% 
  mutate(nutritional_check_new = case_match(
    nutritional_check,
    1 ~ "Never",
    2 ~ "On certain products",
    3 ~ "Very rarely",
    4 ~ "On most products",
    5 ~ "On everything"
  )
  )

exercise_df %>% 
  count(nutritional_check,nutritional_check_new)
```

#### Question 3: Replace Integer respond for veggie consumption with String

```{r}
exercise_df <- rename(exercise_df,"veggie_consumption" = "veggies_day") %>%
  mutate(veggie_consumption_new = case_match(
    veggie_consumption,
    1 ~ "Very unlikely",
    2 ~ "Unlikely",
    3 ~ "Neutral",
    4 ~ "Likely",
    5 ~ "Very likely"
  )
  )

exercise_df %>% 
  count(veggie_consumption, veggie_consumption_new)
```

#### Question 3: Replace Integer respond for fruit consumption with String

```{r}
exercise_df <- rename(exercise_df,"fruit_consumption"= "fruit_day") %>% 
  mutate(fruit_consumption_new = case_match(
    fruit_consumption,
    1 ~ "Very unlikely",
    2 ~ "Unlikely",
    3 ~ "Neutral",
    4 ~ "Likely",
    5 ~ "Very likely"
  )
  )

exercise_df %>% 
  count(fruit_consumption, fruit_consumption_new)
```

## Data Vizualization

#### Question 1:Does exercising regularly affects weight of an individual?
Female Students:
```{r}
#filter Female:
library(ggplot2)
library(here)
exercise_levels <- c("Everyday", "2-3 times a week","Once a week")
weight_female <- weight_df %>% 
  filter(Gender_new == "Female")
ggplot(data=weight_female)+
  geom_bar(mapping= aes(x= factor(exercise_new,exercise_levels), fill=weight_range))+ labs(title="Exercise VS Weight (lbs)", x= "Exercise", y= "Frequency", subtitle= "Female")

ggsave(here("Figures","Exercise VS Weight- Female.png"))
```
Male students:
```{r}
#filter Male:
weight_male <- weight_df %>% 
  filter(Gender_new == "Male")
ggplot(data=weight_male)+
  geom_bar(mapping= aes(x=factor(exercise_new,exercise_levels), fill=weight_range))+ labs(title="Exercise VS Weight (lbs)", x= "Exercise", y= "Frequency", subtitle= "Male")
ggsave(here("Figures","Exercise VS Weight- Male.png"))
```
#### Question 2: Eating out VS Employment status

```{r}
eating_out_levels <- c("Everyday","3-4 times","2-3 times","1-2 times","Never")
ggplot(data=frequency_df)+
  geom_bar(mapping=aes (x=factor(eating_out_new,eating_out_levels), fill=employment_new)) + labs(title = "Frequency of Eating Out VS Employment", x ="Eating out", y="Frequency") + theme(axis.text.x = element_text(angle = 45))

ggsave(here("Figures","Eating out VS Employment.png"))
```
#### Question 2: Eating out VS Income

```{r}
ggplot(data=frequency_df)+
  geom_bar(mapping=aes (x=factor(eating_out_new,eating_out_levels), fill=income_range)) + labs(title = "Frequency of Eating Out VS Income", x ="Eating out", y="Frequency") + theme(axis.text.x = element_text(angle = 45))

ggsave(here("Figures","Eating out VS Income.png"))
```
#### Question 2: Eating out VS Living on or off campus

```{r}
ggplot(data=frequency_df)+
  geom_bar(mapping=aes (x=factor(eating_out_new,eating_out_levels), fill=living_on_off_campus)) + labs(title = "Frequency of Eating Out VS Living on or \noff Campus", x ="Eating out", y="Frequency") + theme(axis.text.x = element_text(angle = 45))
ggsave(here("Figures","Eating out VSLiving on or off campus.png"))
```
#### Question 3: Exercise VS Nutritional check

```{r}
nutri_levels <- c("Never","Very rarely","On certain products", "On most products", "On everything")

ggplot(data=exercise_df)+
  geom_bar(mapping=aes(x= factor(nutritional_check_new, nutri_levels),fill= exercise_new)) + labs(title="Exercise VS Nutritional checks", x= "Nutritional checks", y="Frequency") + theme(axis.text.x = element_text(angle = 45))

ggsave(here("Figures","Exercise VS Nutritional checks.png"))

```
#### Question 3: Exercise VS Veggie consumption

```{r}
Veggie_consumption_levels <- c("Very likely","Likely","Neutral","Unlikely","Very unlikely")
ggplot(data=exercise_df)+ 
  geom_bar(mapping=aes(x= factor(veggie_consumption_new,Veggie_consumption_levels),fill=exercise_new)) + labs(title="Exercise VS Veggie consumption", x= "Veggie consumption", y="Frequency") + theme(axis.text.x = element_text(angle = 45))

ggsave(here("Figures","Exercise VS Veggie consumption.png"))
```


#### Question 3: Exercise VS Fruit consumption

```{r}
Fruit_consumption_levels <- c("Very likely","Likely","Neutral","Unlikely","Very unlikely")
ggplot(data=exercise_df)+
  geom_bar(mapping=aes(fill=exercise_new,x=factor(fruit_consumption_new,Fruit_consumption_levels))) +
  labs(title="Exercise VS Fruit consumption", x= "Fruit consumption", y="Frequency") + theme(axis.text.x = element_text(angle = 45))

ggsave(here("Figures","Exercise VS Fruit consumption.png"))
```

## Basic Statistic test

#### Question 1: Does exercising regularly affects weight of an individual?
Test of independence using Fisher's extract test. alpha value = 0.05
H0: exercise frequency and weight among female college students are independent
Since p = 0.6322 which is more than alpha value (0.05), we accept H0 and can conclude that exercise frequency does not affect weight of female college students in our study.
```{r}
weight_female %>% 
  select("exercise_new","weight_range") %>% 
  table() %>% 
  fisher.test()
```

Since p = 0.7479 which is more than alpha value (0.05), we accept H0 and can conclude that exercise frequency does not affect weight of male college students in our study.
```{r}
weight_male %>% 
  select("exercise_new","weight_range") %>% 
  table() %>% 
  fisher.test()
```

#### Question 2: Does factors such as: Employment, Income, and Living on or off campus affect the frequency of eating out for college students?
Test of independence using Fisher's extract test. alpha value = 0.05
H0: Factor(employment/income/living on or off campus) has no statistical influence to the frequency of college student eating out.
Since p = 0.6443 which is more than alpha value (0.05), we accept H0 and can conclude that employment status of a college student does not affect the frequency of eating out.

Employment:
```{r}
frequency_df %>% 
  select("employment_new","eating_out_new") %>% 
  table() %>% 
  fisher.test()
```

Income:
Since p = 0.6681 which is more than alpha value (0.05), we accept H0 and can conclude that income of a college student does not affect the frequency of eating out.
```{r}
frequency_df %>% 
  select("income_range","eating_out_new") %>% 
  table() %>% 
  fisher.test(input_table, workspace = 1e7) #increase workspace due to error message
```

Living on and off campus:
Since p = 0.2007 which is more than alpha value (0.05), we accept H0 and can conclude that living on or off campus does not affect the frequency of eating out.
```{r}
frequency_df %>% 
  select("living_on_off_campus","eating_out_new") %>% 
  table() %>% 
  fisher.test() 
```

#### Question 3: Does students who exercise regularly more likely to make better food choices?
Test of independence using Fisher's extract test. alpha value = 0.05
H0: Students who exercise regularly are more likely to make better food choices.
Since p = 0.1133 which is more than alpha value (0.05), we accept H0 and can conclude that students who exercise regularly does not check nutritional labels of food more regularly.
```{r}
exercise_df %>% 
  select("nutritional_check_new","exercise_new") %>% 
  table() %>% 
  fisher.test()

```
Since p = 0.1485 which is more than alpha value (0.05), we accept H0 and can conclude that students who exercise regularly does always consumer more vegetables.
```{r}
exercise_df %>% 
  select("veggie_consumption_new","exercise_new") %>% 
  table() %>% 
  fisher.test()
```
Since p = 0.03453 which is less than alpha value (0.05), we reject H0 and can conclude that students who exercise regularly does consumer more fruits.
```{r}
exercise_df %>% 
  select("fruit_consumption_new","exercise_new") %>% 
  table() %>% 
  fisher.test()
```

